Cognate beginnings to bilingual lexical acquisition

### 2023-06-28 docs/index.qmd
Update PDF format

## Participants

```{r}
#| label: fig-participants-time
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 10 
#| fig-cap: "Date at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1, colour = clrs[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
    theme(legend.position = "top",
          legend.title = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 20, hjust = 1),
          axis.title.x = element_blank()) +

    participants %>% 
    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
    group_by(time_stamp, age) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(age) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = age)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
    scale_y_continuous(breaks = seq(0, 500, 50)) +
    scale_colour_manual(values = clrs) +

    plot_layout(ncol = 1) &

    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 20, hjust = 1),
        axis.title.x = element_blank(),
    )
```

### Age

Number of months elapsed between participants' birth date and questionnaire completion.

```{r}
#| label: fig-participants-age
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = "grey60") +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(x = "Age (months)",
         y = "# participants") +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55),
                       breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

### Language exposure (Degree of Exposure)

Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.

```{r fig-participants}
#| label: tab-participants
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:7) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

### SES/parental education

```{r}
#| label: tbl-participants-edu
#| tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
participants %>%
    mutate(id = as.character(id)) |> 
    left_join(bvq_data$logs,
              by = join_by(id, time, age, lp, doe_catalan, doe_spanish, edu_parent)) %>%
    mutate(
        edu_parent = edu_parent |> 
            as.factor() |> 
            fct_na_value_to_level(),
        doe_catalan = cut(doe_catalan,
                          breaks = seq(0, 1, 0.2),
                          include.lowest = TRUE)
    ) %>%
    count(doe_catalan, edu_parent) %>%
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(bvq_data$logs$edu_parent)
        ),
        by = join_by(doe_catalan, edu_parent)
    ) %>%
    rename(total = n) %>%
    pivot_wider(names_from = edu_parent,
                values_from = total,
                values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
    rename(Total = total) %>%
    select(-"NA") %>%
    gt() |> 
    cols_label(doe_catalan = "DoE Catalan") %>%
    sub_missing(columns = everything(),
                missing_text = "--") %>%
    tab_spanner("Educational attainment", 
                Secondary:Primary) %>%
    tab_style(
        style = list(cell_borders(weight = px(2),
                                  sides = "left", 
                                  color = "grey50")),
        locations = cells_body(columns = 8,
                               rows = 1:5)
    ) %>%
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

## Items

```{r}
#| label: tbl-items-summary
#| tbl-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))

items %>%
    unnest(list) %>%
    summarise(n = n(),
              across(c(freq, n_phon, lv),
                     lst(mean, sd, min, max)),
              .by = c(language, list)) |> 
    gt(groupname_col = "language", 
       rowname_col = "list") %>%
    fmt_number(-c(list, n),
               decimals = 1) %>%
    tab_spanner("Frequency (Zipf)",
                columns = starts_with("freq_")) %>%
    tab_spanner("Length (Phonemes)",
                columns = starts_with("n_phon_")) %>%
    tab_spanner("Levenshtein (cognateness)", 
                columns = starts_with("lv_")) %>%
    cols_merge_uncert(col_val = freq_mean,
                      col_uncert = freq_sd) %>%
    cols_merge_uncert(col_val = n_phon_mean,
                      col_uncert = n_phon_sd) %>%
    cols_merge_uncert(col_val = lv_mean, 
                      col_uncert = lv_sd) %>%
    cols_merge_range(col_begin = freq_min, 
                     col_end = freq_max) %>%
    cols_merge_range(col_begin = n_phon_min, 
                     col_end = n_phon_max) %>%
    cols_merge_range(col_begin = lv_min,
                     col_end = lv_max) %>%
    cols_label(list = "",
               n = md("*N*"),
               freq_mean = md("Mean ± SD"),
               freq_min = md("Range"),
               n_phon_mean = md("Mean ± SD"),
               n_phon_min = md("Range"),
               lv_mean = md("Mean ± SD"),
               lv_min = md("Range")) %>%
    tab_style(cell_text(style = "italic"),
              cells_column_labels(everything())) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners(everything()))
```

### Lexical frequency

Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human]. Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [`childesr`](https://langcog.github.io/childes-db-website/) R package. Available words in the corpora were then mapped to their Catalan and Spanish translations. Responses to words with missing lexical frequencies were excluded from analyses.

```{r}
#| label: fig-items-freq
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Distribution of lexical frequency across items. Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations. Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency. The black dot and interval indicate the mean and standard deviation."
mean_sd <- function(x, ...) {
    m <- mean(x, ...)
    s <- sd(x, ...)
    data.frame(mean = m, ymin = m-s, ymax = m+s)
}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = freq)) +
    geom_histogram(fill = "grey60",
                   colour = "white",
                   bins = 30)  +
    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
    )
```

### Word length

Word length ($Length$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.

```{r}
#| label: fig-items-nphon
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>%
    ggplot(aes(x = n_phon, y = n)) +
    facet_wrap(~language, ncol = 1) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = n),
              position = position_nudge(y = 5),
              colour = "black") +
    labs(x = "Phonemes",
         y = "Number of words",
         fill = "Language",
         shape = "Language",
         colour = "Language") +
    scale_x_continuous(breaks = 1:15) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(),
          legend.title = element_blank())
```

### Levenshtein (cognateness)

Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary]. This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1. This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018vocabulary; @fourtassi2020growth; @laing2022phonological for similar approaches].

```{r}
#| label: fig-items-lv
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               scales::percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

### Frequency by word length

```{r}
#| label: fig-items-frequency-phonemes
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2,
               alpha = 0.5,
               colour = "grey60", 
               shape = 1, 
               stroke = 1,
               position = position_jitter(width = 0.1)) +
    geom_smooth(method = "lm",
                formula = "y ~ x" ,
                colour = "black") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

### Frequency by Levenshtein

```{r}
#| label: fig-items-frequency-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

### Word length by Levenshtein

```{r}
#| label: fig-items-nphon-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

### Word exposure index by lexical frequency

We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to. the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).

$$
\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
$$ {#eq-exposure}

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):

$$
\begin{aligned}
\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
\end{aligned}
$$ {#eq-exposure-example}

```{r}
#| label: fig-exposure-distribution
#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
#| warnings: false
#| fig-height: 7
#| fig-width: 13
responses %>%
    mutate(
        exposure = cut(exposure,
                       breaks = seq(0, 9, 1),
                       include.lowest = TRUE)) %>%
    count(exposure, language) %>%
    ggplot(aes(exposure, n)) +
    facet_wrap( ~ language) +
    geom_col(colour = "white",
             fill = "grey60") +
    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
         y = "Number of observations",
         fill = "Language") +

    responses %>%
    mutate(exposure = cut(exposure,
                          breaks = seq(0, 9, 1),
                          include.lowest = TRUE),
           freq = cut(freq, 
                      breaks = seq(0, 9, 0.25), 
                      include.lowest = TRUE)) %>%
    count(exposure, freq, language) %>%
    ggplot(aes(x = freq,
               y = exposure,
               size = n)) +
    facet_wrap( ~ language) +
    geom_point(colour = "grey60",
               alpha = 1/2) +
    labs(x = "Frequency (Zipf score)",
         y = "Exposure-weighted frequency\\
         (Frequency \u00d7 DoE)",
         colour = "Language",
         fill = "Language",
         size = "Number of observations") +
    theme(panel.grid = element_blank()) +
    plot_layout(ncol = 1) &
    scale_colour_manual(values = clrs[c(1, 4)]) &
    scale_fill_manual(values = clrs[c(1, 4)]) &
    guides(fill = "none", colour = "none") &
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = "none")
```

## Data analysis

We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function. The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood (see @fig-likelihood for a graphical representation).

$$
\begin{aligned}
\textbf{Likelihood:} \\
y_{ij} &\sim \text{Cumulative}(p_{k})
\end{aligned}
$$ {#eq-likelihood}

where:

- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
- $i$ is the participant index
- $j$ is the translation equivalent (TE) index
- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.

```{r}
#| label: fig-likelihood
#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire. Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution. For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*. *Familiarity* describes a given participant's familiarity with a given word. When familiarity is low (below some threshold 1), the reulting response is *No*. When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension). When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production). These thresholds are estimated in the cumulative regression model as intercepts 1, and 2. The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely. See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
#| fig-width: 7
#| fig-height: 3.5
dist_df <- tribble(~mean, ~sd, 0,   1)

thresholds <- data.frame(
    ymin = c(-Inf, -1.5, 1.5),
    ymax = c(-1.5, 1.5, Inf),
    xmin = -Inf,
    xmax = Inf,
    .category = str_wrap(c(
        "No", 
        "Understands",
        "Understands and Says"
    ),
    width = 12)
)

thresholds %>%
    ggplot(aes(ymin = ymin, ymax = ymax,
               xmin = xmin, xmax = xmax,
               fill = .category)) +
    geom_rect(alpha = 0.5) +
    stat_dist_slab(
        data = dist_df, 
        aes(dist = "norm", arg1 = mean, arg2 = sd),
        position = "dodge", inherit.aes = FALSE,
        fill = NA, colour = "black"
    ) +
    geom_hline(yintercept = c(-1.5, 1.5), 
               linetype = "dotted",
               linewidth = 1) +
    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
              vjust = 1, hjust = 0.5) +
    annotate(
        geom = "text",
        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
        y = c(-1.5, 2), parse = TRUE,
        x = 0.5, angle = 90, vjust = -0.5
    ) +
    labs(y = "Hypothetical response-generating latent distribution",
         x = "Response probability",
         fill = "Response type",
         colour = "Response type") +
    coord_flip() +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(limits = c(-4, 4)) +
    theme(panel.grid = element_blank(),
          legend.position = "none")
```

To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$. We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression]. For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].

$$
\begin{aligned}
\textbf{Linear model:} \\
logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
\end{aligned}
$$ {#eq-linear}

where:

- $i$ and $j$ index the participant and translation equivalent (TE)
- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively

We used the Bayesian framework to estimate the parameters in our model. This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical]. This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations. We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].

$$
\begin{aligned}
\\
\textbf{Prior:} \\
\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
&&\mbox{[Participant-level coefficient variability]} \\ \\
\Bigg(\begin{smallmatrix}
u_{k_{0}} \\ 
u_{1_{i}} \\ 
\vdots \\ 
u_{8_{i}} 
\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
\Bigg(\Bigg(\begin{smallmatrix}0 \\
0 \\ 
\vdots \\
0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
\vdots & \vdots & \vdots & \vdots \\
\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{u} &\sim LKJcorr(2) \\
\\
&&\mbox{[TE-level coefficient variability]} \\ \\
\Bigg(\begin{smallmatrix}
w_{k_{0}}\\ 
w_{1_{j}} \\ 
\vdots \\ 
w_{3_{j}} 
\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
0\\ 
0 \\ 
\vdots \\
0 
\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
\vdots & \vdots & \vdots & \vdots \\
\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{w_{0-3}} &\sim LKJcorr(2)
\end{aligned}
$$ {#eq-prior}

where:

- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].

::: {.panel-tabset}

#### Prior-predicted mean

```{r fig_prior-mean}
#| label: fig-prior-mean
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| out-width: 90%
#| fig-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

prior_epreds <- add_epred_draws(nd, 
                                model_fit_prior,
                                ndraws = 50, 
                                re_formula = NA) |> 
    filter(.category != "No") |> 
    pivot_wider(names_from = ".category", values_from = ".epred") |> 
    mutate(Understands = Understands + `Understands and Says`) |> 
    pivot_longer(c(Understands, `Understands and Says`),
                 names_to = ".category",
                 values_to = ".epred") |> 
    mutate(.category = factor(.category,
                              levels = c("Understands",
                                         "Understands and Says"),
                              labels = c("Comprehension\n(Understands)",
                                         "Production\n(Understands and Says)")))

prior_epreds |> 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~ .category) +
    geom_line(linewidth = 3/4,
              alpha = 1/2,
              aes(group = .draw),
              colour = "grey50") + 
    labs(x = "Age (months)", 
         y = "Probability of acquisition",
         colour = "Credible interval", 
         fill = "Credible interval") +
    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
                                      mean(responses$age), 
                                      sd(responses$age))[, 1],
                       labels = seq(10, 36, 4)) +
    theme(legend.position = "top",
          axis.ticks.x = element_line(colour = "black"),
          panel.grid = element_blank())
```

#### Prior-predictions

```{r}
#| label: fig-prior-predictions
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

m <- add_predicted_draws(nd, model_fit_prior, ndraws = NULL, re_formula = NA)

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`),
                 names_to = ".prediction", 
                 values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

:::

### Model settings

-   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
-   R version: `r sessionInfo()$R.version$version.string`
-   Algorithm: Hamiltonian Monte Carlo (No U-turn Sampler)
-   Engine: Stan (`brms` interface, `r sessionInfo()$otherPkgs$brms$Version`)
-   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
-   Chains: `r dim(model_fit$fit)[2]`
-   Cores: `r dim(model_fit$fit)[2]`
-   Iterations: `r format(dim(model_fit$fit)[1], big.mark = ",")` (`r format(dim(model_fit$fit)[1]/2, big.mark = ",")`)

More details in Appendix: Session Info.

### R code (`brms`)

```{r}
print(model_fit$formula)
```
:::

# Results

## Descriptive statistics

::: panel-tabset

### By number of phonemes

```{r}
#| lab: fig-responses-age-phonemes
#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
responses %>% 
    mutate(
        age = floor(age),
        n_phon = cut_interval(n_phon, n = 4)
    ) %>%
    count(age, n_phon, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~n_phon) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

### By exposure

```{r}
#| lab: fig-responses-age-exposure
#| fig-cap: "Proportion of responses to each category by age and exposure."
responses %>% 
    mutate(
        age = floor(age),
        exposure = cut_interval(exposure, n = 4)
    ) %>%
    count(age, exposure, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~exposure) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

### By Levenshtein

>>>>

# Data

# Outputs

# Session info
